---
title: "Assignment 2, Q5"
subtitle: "Yun Kyaw, 20177325"
output: pdf_document
date: "2023-02-22"
---

```{r setup, include = FALSE}
library(readr)
library(glmnet)
library(tidyverse)
Q4A2_test <- read_csv("Q4A2_test.csv")
Q4A2_train <- read_csv("Q4A2_train.csv")
```

### A) Perform 10-fold cross-validation (CV) for the Lasso on the training data using the function "cv.glmnet", and plot the CV errors.

```{r}
# let newData be Q4A2_train with dummy variables instead of categorical columns
# (as CV will not work with categorical varaibles)
newData = Q4A2_train %>% 
  mutate(dummy = 1) %>%
  spread(key = League, value = dummy, fill = 0) %>%
  mutate(dummy = 1) %>%
  spread(key = Division, value = dummy, fill = 0) %>%
  mutate(dummy = 1) %>%
  spread(key = NewLeague, value = dummy, fill = 0)

str(newData)

lasso.cv.out = cv.glmnet(as.matrix(newData[,-17]), newData$Salary,
                         lambda = exp(seq(-10, 10, length=1000)), alpha = 1)
plot(lasso.cv.out)
```

### B) Report the model selected by Lasso with $\lambda$ chosen by the "one-standard error" rule.

```{r}
Ytest.1se = predict(lasso.cv.out, s = lasso.cv.out$lambda.1se, newx=as.matrix(newData[,-17]))
```

The model selected by Lasso with lambda chosen by the "one-standard error" rule is:

```{r echo = FALSE}
Ytest.1se
```

The mean square error is:

```{r}
mean((Ytest.1se - newData$Salary)^2)
```

### C) Report the mean square error of the Lasso on the test data with $\lambda$ chosen to be the one minimizing the average CV error.

```{r}
Ytest.min = predict(lasso.cv.out, s = lasso.cv.out$lambda.min, newx=as.matrix(newData[,-17]))
```

The mean square error is:

```{r}
mean((Ytest.min - newData$Salary)^2)
```
