---
title: "STAT 457, A4Q1"
subtitle: Yun Kyaw, 20177325
date: "2023-04-05"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(glmnet)
library(doParallel)

A4Q1_X <- read_csv("A4Q1_X.csv")
A4Q1_Y <- read_csv("A4Q1_Y.csv")
```

### Consider the "biopsy" dataset from the MASS package in R; for its information, load the package and type "? biopsy" in the command line. After removing observations with missing values, the design matrix X is given is given in the file A4Q1_X.csv and the response vector Y in the file A4Q1_Y.csv . Specifically, X has n = 683 rows and 9 columns, and entries of Y take value either "benign" or "malignant".

#### A) Perform PCA on X. Plot the explained variance ratio of each principal component. Find the smallest number of components such that the sum of their ratio is above 90%.

```{r}
# performing PCA
tr_pca = prcomp(A4Q1_X)
tr_pca
```

```{r}
# graphing the explained variance ratio of each principal component
par(mfrow=c(1,2))
ex_var = tr_pca$sdev^2/sum(tr_pca$sdev^2)
plot(ex_var, ylab = 'Explained percentage')
plot(cumsum(ex_var), ylab = 'Cumlative')
```

```{r}
cumsum(ex_var)
```

Thus, we find that the smallest number of components such that the sum of their ratio is above 90% is when there are 5 components.

#### B) For the n observations, consider the first two principal components, and plot each observation as a circle in a 2D plane. For those observations with the response "benign", show their circles in red.

```{r}
Y = A4Q1_Y
par(mfrow=c(1,1))
plot(tr_pca$x[Y == "benign", 1], tr_pca$x[Y == "benign", 2],
     col = 'red', xlim = c(-20, 7), ylim = c(-10, 15),
     xlab ='PC 1', ylab = 'PC 2')
points(tr_pca$x[Y == "malignant", 1], tr_pca$x[Y == "malignant", 2],col = 'blue')
```
